Analysis of Open Data and Computational Reproducibility in Registered Reports in Psychology

### 2019-09-09 reproducing_registered_reports.Rmd
changes through COI statement

While trying to reproduce the results of SPSS scripts, the biggest issue was the often confusing and unclear structure of the scripts. Large portion of the scripts were not annotated, and it was unclear which results they should produce. Often, the descriptive, confirmatory and exploratory analyses were not easily distinguishable because of an overall lack of structure. The absence of understandable variable and value labels in more than half of all SPSS scripts hindered our attempts to reproduce these results. Often the only time-efficient way to check if an article was reproducible so was to run the whole script, and search try to identify specific *p*-values or effect sizes from the article in the SPSS output. SPSS users should take greater care to clearly organize their analysis scripts by adding comments or a README file that links results generated by the SPSS script to the analyses reported in the manuscript. Another frequent problem consisted of missing or incorrectly labeled variables in the dataset, so the script could not run properly. We expect this is the result of authors updating or modifying either their datasets or their scripts during the publication process. This issue could be easily detected if a second author of the manuscript attempted to reproduce the analyses reported in the final manuscript before the data and scripts are shared publicly.

@@

We limited our analysis to Registered Reports based on the idea that these article formats might be used by people who are early adopters of innovations in science, and would therefore be more likely to also share data and code. The default rate with which data and code was shared was relatively high, compared to other studies on the reproducibility of analyses (e.g, Hardwicke et al., 2018, Stockemer et al., 2018), but we do not have data that gives insights into the motivations of these authors. Registered Reports are written by a diverse set of researchers, working in different subfields in psychology, and it would be interesting for future research to qualitatively examine the motivations of researchers who published Registered Reports for sharing or not sharing data and code. There are several good reasons why data can not be shared, and researchers should be able to explain why data is not available when they publish a scientific article [@morey_peer_2016].

The main aim of this article was not to accurately estimate reproducibility rates, but to see what current standards are, and how the reproducibility of research articles using the Registered Report format could be improved. The sample size is small, and it is doubtful whether an accurate estimate of the reproducibility of Registered Reports is of much value, beyond examining where most room for improvement is. Data and code sharing are relatively new, researchers typically lack training in reproducible data analysis, and therefore the main contribution of this article is the identification of common issues that can be improved. We provided some suggestions and examples of better practices that should make the results in published articles more reproducible.

In addition to the recommendations we have provided above, we believe novel technological solutions might improve the reproducibility of research articles. For example, Code Ocean is an online, cloud based, computational reproducibility platform [@clyburne-sherin_computational_2018]. It provides a code environment (or container) that runs online, which means that researchers using Code Ocean do not have to download data, code, or software, but can analyze the data in their browser. It is not yet possible to use SPSS within Code Ocean, but for R code it solves the problem of package versions (since the container uses the version of packages specified by the researchers) and file locations.^[For a Code Ocean capsule reproducing this manuscript: https://codeocean.com/capsule/8580632/tree/v1] Other platforms in the reproducibility space include Whole Tale [@brinckman2019computing], "a research environment that captures and, at the time of publication, exposes salient details of the entire research process via access to persistent versions of the data and code used, provenance, and data lineage" (p. 855); and Binder [@ragan2018binder], an open-source, browser-based tool for creating and sharing reproducible environments.

Another useful technology is RMarkdown, which enable researchers to write fully executable manuscripts. RMarkdown files load the raw data and allow researchers to compute each number reported in the manuscript from the data, instead of copy-pasting values. This means that, as long as the data and require packages can be loaded, all reported numbers can be reproduced. This saves a lot of time matching the analysis code to the reported results, and thus speeds up the process of checking whether all results reported in the manuscripts are reproduced by the code. The current manuscript is an example of a reproducible RMarkdown file.^[See https://github.com/Lakens/reproducing_registered_reports/blob/master/reproducing_registered_reports.Rmd] Additional solutions that help researchers to share reproducible analyses might become available in the future.

Finally, journals who value reproducibility might find it worthwhile to check whether the data and analysis code shared with a submission can be used to reproduce the results. The average time it took our team to check that the analysis code could reproduce the results reported in the paper was `r round(average_time,0)` minutes. This is slightly shorter than the time it took Hardwicke and colleagues (2018), who estimated (without keeping track of the time explicitly) that reproducibility reports took between 2 and 25 person-hours, depending on whether the paper eventually fell in the reproducible or not-reproducible category, and whether author assistance was needed. The main difference might be that we did not write 'reproducibility reports' for each article. Documenting the process of reproducing a paper adds transparency and allows others to check the the decisions about every value in an article. Whether such a level of detail is worth the additional time investment of documenting each reported value is a cost-benefit analysis that journals should make for themselves. The required time might be reduced by explicitly asking authors to submit files in a format or structure that facilitates such checks, or automating part of the work that is needed to check the reproducibility of results. Overall, we feel that the time required for a basic check of the reproducibility of manuscripts (i.e., where one checks whether the *main* results in the paper are reproduced by the analysis scripts, but without documenting this step at the level of each individual number) is not an insurmountable problem for scientific journals that aim to add value, and would substantially improve the computational reproducibility of the published literature.

In addition to novel technologies, most progress can probably be made by developing standards within a research community, and educating researchers about best practices that guarantee reproducibility. Most researchers are not trained in reproducible data analysis, and can not be expected to invent best practices from scratch. As good examples appear in the published literature over time, and best practices within subdisciplines crystalize, standards should emerge that improve reproducibility, and that allow researchers to share data and code in such a way that others with basic scientific training can reproduce their results and reuse their data.

>>>>

While trying to reproduce the results of SPSS scripts, the biggest issue was the often confusing and unclear structure of the scripts. Large portion of the scripts were not annotated, and it was unclear which results they should produce. Often, the descriptive, confirmatory and exploratory analyses were not easily distinguishable because of an overall lack of structure. The absence of understandable variable and value labels in more than half of all SPSS scripts hindered our attempts to reproduce these results. Often the only time-efficient way to check if an article was reproducible was to run the whole script, and try to identify specific *p*-values or effect sizes from the article in the SPSS output. SPSS users should take care to clearly organize their analysis scripts by adding comments or a README file that links results generated by the SPSS script to the analyses reported in the manuscript. Another frequent problem was missing or incorrectly labeled variables in the dataset, so the scripts could not run properly. We expect this is the result of authors updating or modifying either their datasets or their scripts during the publication process. This issue could be easily detected if a second author of the manuscript attempted to reproduce the analyses reported in the final manuscript before data and scripts are shared publicly.

@@

We limited our analysis to Registered Reports based on the idea that these article formats might be used by people who are early adopters of innovations in science, and would therefore be more likely to also share data and code. We found that the rate at  which data and code was shared in our sample was high, relative to studies anlayzed in other reproducibility-focused reviews (e.g, Hardwicke et al., 2018, Stockemer et al., 2018), but we do not have data that gives insights into the motivations of these authors. Registered Reports are written by a diverse set of researchers, working in different subfields in psychology, and it would be interesting for future research to qualitatively examine the motivations of researchers who published Registered Reports for sharing or not sharing data and code. There are several good reasons why some data should not be shared, and when applicable, researchers should be encouraged to explain their reasons [@morey_peer_2016].

The main aim of this article was not to precisely estimate reproducibility rates, but to see what current standards are, and how the reproducibility of research articles using the Registered Report format could be improved. The sample size is small, and it is doubtful whether a precise estimate of the reproducibility of Registered Reports is of much value, beyond examining where most room for improvement is. Data and code sharing are relatively new, researchers typically lack training in reproducible data analysis, and therefore the main contribution of this article is the identification of common issues that can be improved. We provided some suggestions and examples of better practices that should make the results in published articles more reproducible.

In addition to the recommendations we have provided above, we believe novel technological solutions might improve the reproducibility of research articles. For example, Code Ocean is an online, cloud based, computational reproducibility platform [@clyburne-sherin_computational_2018]. It provides a code environment (or container) that runs online, which means that researchers using Code Ocean do not have to download data, code, or software, but can analyze the data in their browser. It is not currently possible to use SPSS within Code Ocean, but for R code, it solves the problem of package versions (since the container uses the version of packages specified by the researchers) and file locations.^[For a Code Ocean capsule reproducing this manuscript, see https://doi.org/10.24433/CO.4275368.v1] Other platforms in the reproducibility space include Whole Tale [@brinckman2019computing], "a research environment that captures and, at the time of publication, exposes salient details of the entire research process via access to persistent versions of the data and code used, provenance, and data lineage" (p. 855); and Binder [@ragan2018binder], an open-source, browser-based tool for creating and sharing reproducible environments.

Another useful technology is RMarkdown, which enable researchers to write fully executable manuscripts. RMarkdown files load the raw data and allow researchers to compute each number reported in the manuscript from the data, instead of copy-pasting values. This means that, as long as the data and require packages can be loaded, all reported numbers can be reproduced. This saves time when matching the analysis code's output to reported results, and thus speeds up the process of checking whether all results reported in the manuscripts are reproduced. The current manuscript is an example of a reproducible RMarkdown file.^[See https://github.com/Lakens/reproducing_registered_reports/blob/master/reproducing_registered_reports.Rmd] Additional solutions that help researchers to share reproducible analyses may become available in the future.

Finally, journals that value reproducibility might find it worthwhile to check whether the data and analysis code shared with a submission can be used to reproduce the results. The average time it took our team to check that the analysis code could reproduce the results reported in the paper was `r round(average_time,0)` minutes. This is slightly shorter than the time it took Hardwicke and colleagues (2018), who estimated (without keeping track of the time explicitly) that reproducibility reports took between 2 and 25 person-hours, depending on whether the paper eventually fell in the reproducible or not-reproducible category, and whether author assistance was needed. One major difference between our approaches is that we did not write 'reproducibility reports' for each article. Documenting the process of reproducing a paper adds transparency and allows others to check the the decisions about every value in an article. Whether such a level of detail is worth the additional time investment of documenting each reported value is a cost-benefit analysis that journals should make for themselves. The required time might be reduced by explicitly asking authors to submit files in a format or structure that facilitates such checks, or automating part of the work that is needed to check the reproducibility of results. Overall, we feel that the time required for a basic check of the reproducibility of manuscripts (i.e., where one checks whether the *main* results in the paper are reproduced by the analysis scripts, but without documenting this step at the level of each individual number) is a surmountable hurdle for journals, and would substantially improve the computational reproducibility of the published literature.

In addition to novel technologies, most progress can probably be made by developing standards within research communities, and educating researchers about best practices that guarantee reproducibility. Most researchers are not trained in reproducible data analysis, and cannot be expected to invent best practices from scratch. As good examples appear in the published literature over time, and best practices within subdisciplines crystalize, standards should emerge that improve reproducibility, and that allow researchers to share data and code in such a way that others with basic scientific training can reproduce their results and reuse their data.

# Conflict of Interest Statement
One author (SG) worked at Code Ocean during the writing of this manuscript. SG joined the project after its accompanying Code Ocean component was already submitted and published, and did not write any of the text concerning Code Ocean.
